---
title: "Exercises"
author: "Pedro Ojeda, Birgitte Brydsö, Mirko Myllykoski, Lars Viklund"
date: "May 27th, 2020"
output: html_document
---

## Problem 1.

When the *user* time is longer than the *elapsed* time from system.time() function?

## Problem 2.

In this problem, you should use the templates to write your own code. 

Given the matrix A of ones with a size of 5000x5000:

```{r eval=FALSE}
A <- #initialize the matrix
```

compare the profiling results of the following functions:

* a) a function sumcol to compute the sum of the elements by columns, you can use the following template,


```{r eval=FALSE}
sumcol <- function(B) {
   l <- #obtain the number of columns
   colsm <- #create a vector to save the sums

   for(i in 1:l){  
      s <- 0  #collect partial sums
      for(j in 1:l){
          s <- s + #sum the columns of the matrix
      }
      colsm[i] <- s
   }
   return(colsm)
}
res1 <- sumcol(A)
```


* b) the R built-in function

```{r eval=FALSE}
colSums()
```


* Use the Rprof package to obtain the profiling summary. 
* Are the performances of the two functions similar? 
* The two functions do the same calculation, why the performaces could differ?
* What is the most time consuming part in the function **sumcol()**? *Hint: One can use tic/toc functions for monitoring parts of the functions.*
* **Challenge:** Obtain the profiling summary by using microbenchmark and rbenchmark packages.