---
title: "Solution"
output: html_document
---

## Problem 1. 

The first function sumcol uses 3.8Mb of memory:

```{r eval=FALSE}

A <- matrix(1.0, 5000, 5000)
sumcol <- function(B) {
  l <- ncol(B)  #obtain the number of columns
  colsm <- rep(0,l) #create a vector to save the sums
  
  for(j in 1:l){  
    s <- 0
    for(i in 1:l){
      s <- s + B[i,j]
    }
    colsm[i] <- s
  }
  return(colsm)
}

Rprof("Rprof-mema.out", memory.profiling=TRUE)
res1 <- sumcol(A)
Rprof(NULL)
summaryRprof("Rprof-mema.out", memory="both")

$by.self
                 self.time self.pct total.time total.pct
"sumcol"              1.60    98.77       1.62    100.00
"getAssignedVar"      0.02     1.23       0.02      1.23
                 mem.total
"sumcol"               3.8
"getAssignedVar"       0.0

$by.total
                       total.time total.pct mem.total self.time
"sumcol"                     1.62    100.00       3.8      1.60
"getAssignedVar"             0.02      1.23       0.0      0.02
"addVar"                     0.02      1.23       0.0      0.00
"assign"                     0.02      1.23       0.0      0.00
"cmpfun"                     0.02      1.23       0.0      0.00
"compiler:::tryCmpfun"       0.02      1.23       0.0      0.00
"doTryCatch"                 0.02      1.23       0.0      0.00
"findLocals1"                0.02      1.23       0.0      0.00
"findLocalsList"             0.02      1.23       0.0      0.00
"findLocalsList1"            0.02      1.23       0.0      0.00
"FUN"                        0.02      1.23       0.0      0.00
"funEnv"                     0.02      1.23       0.0      0.00
"lapply"                     0.02      1.23       0.0      0.00
"make.functionContext"       0.02      1.23       0.0      0.00
"tryCatch"                   0.02      1.23       0.0      0.00
"tryCatchList"               0.02      1.23       0.0      0.00
"tryCatchOne"                0.02      1.23       0.0      0.00
                       self.pct
"sumcol"                  98.77
"getAssignedVar"           1.23
"addVar"                   0.00
"assign"                   0.00
"cmpfun"                   0.00
"compiler:::tryCmpfun"     0.00
"doTryCatch"               0.00
"findLocals1"              0.00
"findLocalsList"           0.00
"findLocalsList1"          0.00
"FUN"                      0.00
"funEnv"                   0.00
"lapply"                   0.00
"make.functionContext"     0.00
"tryCatch"                 0.00
"tryCatchList"             0.00
"tryCatchOne"              0.00

$sample.interval
[1] 0.02

$sampling.time
[1] 1.62
```

The second function allocates 0.1Mb memory:

```{r eval=FALSE}
Rprof("Rprof-memb.out", memory.profiling=TRUE)
res2 <- colSums(A)
Rprof(NULL)
summaryRprof("Rprof-memb.out", memory="both")

$by.self
          self.time self.pct total.time total.pct mem.total
"colSums"      0.02      100       0.02       100       0.1

$by.total
          total.time total.pct mem.total self.time self.pct
"colSums"       0.02       100       0.1      0.02      100

$sample.interval
[1] 0.02

$sampling.time
[1] 0.02
```

## Problem 2.

By using the bm array, one can compute the mean and standard deviation as follows:

```{r eval=TRUE}
library(bigmemory)
set.seed(1234)
bm <- big.matrix(1e7, 3, backingfile = "bm", backingpath = getwd())
bm
```

now, work with chunks of $10^6$ rows,

```{r eval=TRUE}
chunksize <- 1e6
start <- 1
while (start <= nrow(bm)) {
  end <- min(start + chunksize -1, nrow(bm))
  chunksize <- end - start + 1
  bm[start:end, 1] <- rpois(chunksize, 1e3)
  bm[start:end, 2] <- sample(0:1, chunksize, TRUE, c(0.7,0.3))
  bm[start:end, 3] <- runif(chunksize, 0, 1e5)
  start <- start + chunksize
}
```

means:

```{r eval=TRUE}
col.sums <- numeric(3)
chunksize <- 1e6
start <- 1
while (start <= nrow(bm)) {
  end <- min(start + chunksize -1, nrow(bm))
  col.sums <- col.sums + colSums(bm[start:end,])
  start <- start + chunksize
}
col.means <- col.sums / nrow(bm)
```


standard deviations:

```{r eval=TRUE}
col.sq.dev <- numeric(3)
chunksize <- 1e6
start <- 1
while (start <= nrow(bm)) {
  end <- min(start + chunksize -1, nrow(bm))
  col.sq.dev <- col.sq.dev + rowSums((t(bm[start:end,]) - col.means) ^2)
  start <- start + chunksize
}
col.var <- col.sq.dev/(nrow(bm)-1)
col.sd <- sqrt(col.var)
```

## Problem 3.

Execution times are:

```{r eval=FALSE}
                           1st. function   
vector size          1e+06     1e+07    1e+08
time(sec)             0.05      0.5      4.89 
                          Vectorized function
time(sec)             0.0       0.1     0.81
```
                          
based on this data, the vectorized function had the best performance. However, one needs to take into account the memory usage which in the case of the vectorized one is at least twice of the initial vector **x**.