---
title: "Exercises"
author: "Pedro Ojeda, Birgitte Brydsö, Mirko Myllykoski, Lars Viklund"
date: "May 27th, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

Take a look at the code for the credit bank data set in the script **parallel.R** and also at the **job_parallel.sh** batch script. Set the number of cores you want to use in both scripts first to 1 and then increase the number up to 6. Upload the working folder to Kebnekaise and run the simulation. Monitor the timings for the simulations. You may use the *job-usage* tool on Kebnekaise.

## Problem 2
 
* Create a matrix A of ones with a size 5000x2000. Analyze the timings and memory usage with tictoc and gcinfo() respectively of the following function:

```{r eval=FALSE}
res1 <- foreach(i=1:nrow(A), .combine='rbind') %dopar% (A[i,]/mean(A[i,]))
```

* What are the timings when one goes from 1 core to 2 cores? Does the performance improve considerably upon increasing the number of cores? 

* What is the memory usage reported by gcinfo()? Could this result be linked to the performance observed when going from 1 to 2 cores?

## Problem 3.

* Compare the performance of the following function for computing the eigenvalues of a matrix of *NxN* by using a different number of cores and using the **foreach** function, with *N* going from 1 to 500:

```{r eval=FALSE}
max.eig <- function(N, sigma) {
    d <- matrix(rnorm(N**2, sd = sigma), nrow = N)

    E <- eigen(d)$values
    
    abs(E)[[1]]
}
```


* How is the performance of the **foreach** function with 1 core compared with that of the **sapply** function (N=500):

```{r eval=FALSE}
res3 <- sapply(1:N, function(n) {max.eig(n, 1)})
```



